Enforced Invariant Confluence

### 2018-04-17 doc/paper/sections/abstract.tex
Minor edits + starting on eval.

\begin{abstract}
Strongly consistent systems are easy to reason about but face fundamental
limitations in availability and performance. Weakly consistent systems can be
implemented with very high performance, but place a burden on the application
developer to reason about complex interleavings of execution.
Invariant-confluence provides a formal framework for understanding when we can
get the best of both worlds. An invariant-confluent object can be efficiently
replicated with no coordination needed to preserve its invariants.  However,
actually determining whether or not an object is invariant-confluent is
challenging (undecidable). In this paper, we establish conditions under which a
commonly used sufficient condition for invariant confluence is both necessary
and sufficient. We use this condition to design a general-purpose interactive
invariant-confluence decision procedure and a novel sufficient condition which
can be checked automatically. We then take a step beyond invariant-confluence
and introduce a generalization of invariant-confluence, called segmented
invariant-confluence, which allows us to replicate non-invariant-confluent
objects with a small and tunable amount of coordination. Implementing our
theoretical findings in a system called Lucy, we evaluate our methods on a
number of case studies and benchmarks showing that... TODO: Perform evaluation
:P.
% Strongly consistent systems are easy to reason about but are slow. Weakly consistent systems are fast but hard to reason about. Invariant-confluence gives a formal framework to reason about when data can be replicated without any coordination needed to maintain invraints, the best of borth worlds. However, determining whether an object is invariant-confluent is hard. In this paper, we establish conditions under which a commonly used sufficient condition for invariant confluence is both necessary and usfficient. We use this condition to design a general-purpose interactive decision procedure and a novel sufficient condition which can be checked automatically. We then take a step beyond invariant-confluence and introduce a generalization of invariant-confuluence, called segmented invariant-confluence which allows to replicate objects with a minimal amount of coordination that are otherwise not invariant-confluent. We implement all this in a system called lucy and look at a bunch of stuff.
\end{abstract}

>>>>

\begin{abstract}
Strongly consistent systems are easy to reason about but face fundamental
limitations in availability and performance, and weakly consistent systems can
be implemented with very high performance but place a burden on the
application developer to reason about complex interleavings of execution.
Invariant-confluence provides a formal framework for understanding when we can
get the best of both worlds. An invariant-confluent object can be efficiently
replicated with no coordination needed to preserve its invariants.  However,
actually determining whether or not an object is invariant-confluent is
challenging (undecidable). In this paper, we establish conditions under which a
commonly used sufficient condition for invariant confluence is both necessary
and sufficient and use this condition to design (a) a general-purpose
interactive invariant-confluence decision procedure and (b) a novel sufficient
condition which can be checked automatically. We then take a step beyond
invariant-confluence and introduce a generalization of invariant-confluence
called segmented invariant-confluence, which allows us to replicate
non-invariant-confluent objects with a small and tunable amount of
coordination. Implementing our theoretical findings in a system called Lucy, we
evaluate our methods on a number of case studies and benchmarks showing that...
TODO: Perform evaluation :P.
% Strongly consistent systems are easy to reason about but are slow. Weakly
% consistent systems are fast but hard to reason about. Invariant-confluence
% gives a formal framework to reason about when data can be replicated without
% any coordination needed to maintain invraints, the best of borth worlds.
% However, determining whether an object is invariant-confluent is hard. In
% this paper, we establish conditions under which a commonly used sufficient
% condition for invariant confluence is both necessary and usfficient. We use
% this condition to design a general-purpose interactive decision procedure and
% a novel sufficient condition which can be checked automatically. We then take
% a step beyond invariant-confluence and introduce a generalization of
% invariant-confuluence, called segmented invariant-confluence which allows to
% replicate objects with a minimal amount of coordination that are otherwise
% not invariant-confluent. We implement all this in a system called lucy and
% look at a bunch of stuff.
\end{abstract}

### 2018-04-17 doc/paper/sections/evaluation.tex
Minor edits + starting on eval.

\section{Evaluation}\seclabel{Evaluation}
\subsection{Lucy}
\todo{Describe implementation of lucy.}

\subsection{Case Studies}
\todo{Bank account with and without decrements.}
\todo{Auction example.}
\todo{Foreign keys with and without deletions.}

\subsection{TPC and Feral}
\todo{See which of the Feral and TPC transactions we can determine automatically}

\subsection{Segmented Invariant Confluence}
\todo{one graph: number of locked operations x axis; throughput on the y-axis}
\todo{one graph: number of segments on the x-axis; throughput on the y-axis}
\todo{compare with no coordination and "VR"}

>>>>

\section{Evaluation}\seclabel{Evaluation}
In this section, we describe Lucy: an implementation of our decision procedures
and system models. We then evaluate Lucy by answering the following questions:
\begin{itemize}
  \item
    How practical is the interactive invariant-confluence decision procedure?
    Can we use it to classify real-world transactions and invariants?
  \item
    How practical is segmented invariant-confluence? Are real-world workloads
    amenable to segmentation?
  \item
    How efficient is the interactive invariant-confluence decision procedure?
  \item
    How efficiently can we replicate a segmented invariant-confluent object as
    compared to alternative approaches like replicating with weak or strong
    consistency?
  \item
    How does the performance of replicating a segmented invariant-confluence
    object vary as we vary the workload, segmentation, and replication factor?
\end{itemize}

\subsection{Lucy}
Lucy includes an implementation of the interactive decision procedure described
in \algoref{InteractiveDecisionProcedure}, an implementation of a decision
procedure which checks criteria (1) - (4) from \thmref{LatticeProperty}, and an
implementation of the decision procedure described in
\algoref{ArbitraryStartInteractiveDecisionProcedure}. The decision procedures
are implemented in roughly 2,500 lines of Python. Users specify objects,
transactions, and invariants in a small Python DSL and interact with the
interactive decision procedures using an interactive Python console. We use
Z3~\cite{de2008z3} to implement our invariant-closure decision procedure,
compiling objects and invariants into a formula which is satisfiable if and
only if the object is \emph{not} invariant-confluent. If the object is
invariant-closed, then Z3 concludes that the formula is unsatisfiable.
Otherwise, if the object is not invariant-closed, then Z3 produces a
counterexample witnessing the satisfiability of the formula.

Lucy also includes an implementation of the invariant-confluence and
segmented-invariant confluence system models in roughly 3,500 lines of C++.
Users specify objects, transactions, invariants, and segmentations in C++. Lucy
then replicates the objects using invariant-confluence, segmented
invariant-confluence, eventual consistency, or linearizability. In
\secref{SegmentedInvariantConfluenceEval}, we compare the performance of these
various consistency models.
%
Because fault-tolerance is largely an orthogonal concern to
invariant-confluence, Lucy is implemented without fault-tolerance. It would be
straightforward to add fault-tolerance to Lucy, but it would not affect our
discussions or evaluation, so we leave it for future work. Moreover, users
currently have to specify their workloads in Python (for the decision
procedures) and C++ (for the runtime). In the future, we plan on removing this
redundancy.

\subsection{Decision Procedures}
% \todo{Bank account with and without decrements.}
% \todo{Auction example.}
% \todo{Foreign keys with and without deletions.}
% \subsection{TPC and Feral}
% \todo{See which of the Feral and TPC transactions we can determine automatically}

\subsection{Segmented Invariant Confluence}\seclabel{SegmentedInvariantConfluenceEval}
% \todo{one graph: number of locked operations x axis; throughput on the y-axis}
% \todo{one graph: number of segments on the x-axis; throughput on the y-axis}
% \todo{compare with no coordination and "VR"}

\begin{figure}[ht]
  \centering

  \begin{subfigure}[c]{\columnwidth}
    \centering
    \includegraphics[width=\columnwidth]{figures/vary_withdraws.pdf}
  \end{subfigure}
  \begin{subfigure}[c]{\columnwidth}
    \includegraphics[width=\columnwidth]{figures/vary_segments.pdf}
  \end{subfigure}

  \caption{}\figlabel{}
\end{figure}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\columnwidth]{figures/vary_nodes.pdf}
  \caption{}\figlabel{}
\end{figure}

\uselengthunit{in}\printlength{\columnwidth}

### 2018-04-17 doc/paper/sections/introduction.tex
Minor edits + starting on eval.

\section{Introduction}
When an application designer decides to replicate a piece of data, they have to
make a fundamental choice between weak and strong consistency. Replicating the
data with strong consistency using a technique like state machine
replication~\cite{schneider1990implementing, lamport1998part,
liskov2012viewstamped, ongaro2014search} makes the application designer's life
very easy. To the application developer, a strongly consistent system behaves
exactly like a single-threaded system running on a single node, so reasoning
about the behavior of the system is simple. Unfortunately, strong consistency
is at odds with performance. The CAP theorem and PACELC theorem tell us that
strongly consistent systems suffer from higher latency at best and
unavailability at worst~\cite{brewer2012cap, abadi2012consistency}. On the
other hand, weak consistency models like eventual~\cite{vogels2009eventually},
PRAM~\cite{lipton1988pram}, and causal~\cite{ahamad1995causal} consistency
allow data to be replicated with high availability and low latency, but they
put a tremendous burden on the application designer to reason about the complex
interleavings of operations that are allowed by these weak consistency models.
In fact, weak consistency models strip an application developer of one of the
earliest and most effective tools that is used to reason about the execution of
programs: application invariants~\cite{hoare1969axiomatic}. Even if every
transaction executing in a weakly consistent system maintains an application
invariant, the system as a whole can produce invariant-violating states.

@@

Is it possible for us to have our strongly consistent cake and eat it with high
availability too? Can we replicate a piece of data with weak consistency but
still ensure that its invariants are maintained? Yes, sometimes. Bailis et al.\
introduced the notion of invariant-confluence as a necessary and sufficient
condition for when invariants can be maintained over replicated data without
the need for any coordination~\cite{bailis2014coordination}. If an object is
invariant-confluent with respect to an invariant, we can replicate it with the
performance benefits of weak consistency and (some of) the correctness benefits
of strong consistency. However, the task of identifying whether or not an
object actually is invariant-confluent remain a challenge. Bailis et al.\
manually categorized a set of common objects, transactions, and invariants
(e.g.\ foreign key constraints on relations, linear constraints on integers) as
invariant-confluent or not, but ideally we would have a general purpose program
that can automatically determine invariant-confluence for us. \textbf{The first
main thrust of this paper is to make progress towards a general-purpose
invariant-confluence decision procedure.}

>>>>

\section{Introduction}
When an application designer decides to replicate a piece of data, they have to
make a fundamental choice between weak and strong consistency. Replicating the
data with strong consistency using a technique like state machine
replication~\cite{schneider1990implementing, lamport1998part,
liskov2012viewstamped, ongaro2014search} makes the application designer's life
very easy. To the application developer, a strongly consistent system behaves
exactly like a single-threaded system running on a single node, so reasoning
about the behavior of the system is simple. Unfortunately, strong consistency
is at odds with performance. The CAP theorem and PACELC theorem tell us that
strongly consistent systems suffer from higher latency at best and
unavailability at worst~\cite{brewer2012cap, abadi2012consistency}. On the
other hand, weak consistency models like eventual~\cite{vogels2009eventually},
PRAM~\cite{lipton1988pram}, and causal~\cite{ahamad1995causal} consistency
allow data to be replicated with high availability and low latency, but they
put a tremendous burden on the application designer to reason about the complex
interleavings of operations that are allowed by these weak consistency models.
In fact, weak consistency models strip an application developer of one of the
earliest and most effective tools that is used to reason about the execution of
programs: application invariants~\cite{hoare1969axiomatic}. Even if every
transaction executing in a weakly consistent system individually maintains an
application invariant, the system as a whole can produce invariant-violating
states.

@@

Is it possible for us to have our strongly consistent cake and eat it with high
availability too? Can we replicate a piece of data with weak consistency but
still ensure that its invariants are maintained? Yes, sometimes. Bailis et al.\
introduced the notion of invariant-confluence as a necessary and sufficient
condition for when invariants can be maintained over replicated data without
the need for any coordination~\cite{bailis2014coordination}. If an object is
invariant-confluent with respect to an invariant, we can replicate it with the
performance benefits of weak consistency and (some of) the correctness benefits
of strong consistency. However, the task of identifying whether or not an
object actually is invariant-confluent remains a challenge. Bailis et al.\
manually categorized a set of common objects, transactions, and invariants
(e.g.\ foreign key constraints on relations, linear constraints on integers) as
invariant-confluent or not, but ideally we would have a general purpose program
that can automatically determine invariant-confluence for us. \textbf{The first
main thrust of this paper is to make progress towards a general-purpose
invariant-confluence decision procedure.}
